### 2018-03-08 paper/background.tex
Add dynamic dependencies

\noindent
The above makefile lists three tasks: (i) compile a utility library comprising
files \cmd{util.h} and \cmd{util.c} into \cmd{main.o} by
executing\footnote{In this example we treat \cmd{gcc} as a pure function for the
sake of simplicity. In reality there are multiple versions of \cmd{gcc} and the
actual binary that is used to compile and link files is also listed as a task
dependency.} the command \cmd{gcc -c util.c}, (ii) compile the main source file
\cmd{main.c} into \cmd{main.o}, and (iii) link object files \cmd{util.o} and
\cmd{main.o} into the executable \cmd{main.exe}. The makefile contains the
complete information about the task \emph{dependency graph}, which is shown in
Fig.~\ref{fig-make}(a).

\begin{figure}[h]
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example.pdf}}
\caption{A dependency graph}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-full.pdf}}
\caption{A full rebuild}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-partial.pdf}}
\caption{A partial rebuild}
\end{subfigure}
\caption{A dependency graph and two build scenarios. Input files are shown as
rectangles, intermediate and output files are shown as rounded rectangles. Dirty
inputs and files that are rebuilt are highlighted.
\label{fig-make}}
\end{figure}

@@

The requirement to execute tasks \emph{at most once} and only if they
\emph{transitively depend on dirty inputs} is essential for build systems,
their raison d'\^etre. We will call build systems that satisfy this requirement
\emph{minimal}.

To achieve minimality \Make relies on two key ideas. First, it uses \emph{file
modification time} to detect the files that are dirty: a file is marked dirty if
it was modified after the previous build. Second, it constructs a complete task
dependency graph from the information contained in the makefile and executes
tasks in the \emph{topological order}.

>>>>

\noindent
The above makefile lists three tasks: (i) compile a utility library comprising
files \cmd{util.h} and \cmd{util.c} into \cmd{main.o} by
executing\footnote{In this example we treat \cmd{gcc} as a pure function for the
sake of simplicity. In reality there are multiple versions of \cmd{gcc} and the
actual binary that is used to compile and link files is also listed as a task
dependency.} the command \cmd{gcc -c util.c}, (ii) compile the main source file
\cmd{main.c} into \cmd{main.o}, and (iii) link object files \cmd{util.o} and
\cmd{main.o} into the executable \cmd{main.exe}. The makefile contains the
complete information about the \emph{task dependency graph}, which is shown in
Fig.~\ref{fig-make}(a).

\begin{figure}[h]
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example.pdf}}
\caption{Task dependency graph}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-full.pdf}}
\caption{Full rebuild}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-partial.pdf}}
\caption{Partial rebuild}
\end{subfigure}
\caption{A task dependency graph and two build scenarios. Input files are shown as
rectangles, intermediate and output files are shown as rounded rectangles. Dirty
inputs and files that are rebuilt are highlighted.
\label{fig-make}}
\end{figure}

@@

The requirement to \emph{execute tasks at most once and only if they transitively
depend on dirty inputs} is essential for build systems, it is their raison
d'\^etre. We will call build systems that satisfy this requirement \emph{minimal}.

To achieve minimality \Make relies on two main ideas. First, it uses \emph{file
modification time} to detect the files that are dirty: a file is marked dirty if
it was modified after the previous build. Second, it constructs a complete task
dependency graph from the information contained in the makefile and executes
tasks in a \emph{topological order}.

Note that it is not always possible to know the dependencies of a source file
upfront, or \emph{statically}. To demonstrate this, let us first add another
task to the above makefile:

\begin{minted}[frame=single]{makefile}
release.tar: main.exe README
    tar -cf release.tar main.exe README
\end{minted}

So far so good: if either \cmd{main.exe} or \cmd{README} is dirty, \Make will
recreate the release archive. But now imagine that we need to specify the files
that go into \cmd{release.tar} in a text file \cmd{release.txt}. The dependencies
of \cmd{release.tar} are no longer known statically: they \emph{depend on the
contents} of \cmd{release.txt}, which might not even exist before the build
starts, e.g. if it is produced by concatenating input files \cmd{release-bins.txt}
and \cmd{release-docs.txt}. Alas, makefiles cannot express such \emph{dynamic
dependencies}, requiring ad hoc workarounds such as \emph{build phases}, which
are problematic~\cite{hadrian}.
